Ms 1

### 2017-09-01 ms1/ms1_display.Rmd
adapting

est2pct = function(x) (exp(x) - 1)*100
model_ci = function(df) {
  df %>%
    mutate(pct_change=est2pct(estimate),
           ci=1.96*std.error,
           ci_low=est2pct(estimate - ci),
           ci_high=est2pct(estimate + ci),
           change_str=sprintf('%.2f%% (%.2f -- %.2f%%)', pct_change, ci_low, ci_high))
}

model_abx = read_tsv('tables/model_abx.tsv') %>%
  rename(antibiotic=name) %>%
  model_ci()

@@

# this should be the same list as for the models
stopifnot(all(top_abx %in% model_abx$antibiotic))

@@

```{r claims_by_age}
change_demog = read_tsv('tables/model_overall_demography.tsv') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  model_ci() %>%
  filter(term=='year0') %>%
  select(name, adj_change=change_str)

@@

```{r adjusted_changes}
adjusted_changes = model_abx %>%
  filter(term=='year0') %>%
  select(antibiotic, pct_change, ci_low, ci_high) %>%
  arrange(pct_change) %>%
  mutate(antibiotic=fct_inorder(antibiotic),
         antibiotic=fct_relevel(antibiotic, 'overall', after=Inf))

unadjusted_changes = bind_rows(claims_by_abx, claims_overall) %>%
  filter(antibiotic %in% c('overall', top_abx), year %in% c(2011, 2014)) %>%
  select(antibiotic, year, cpkp) %>%
  spread(year, cpkp) %>%
  mutate(change=(`2014`-`2011`)/3,
         fold_change=change/`2011`,
         pct_change=100*fold_change) %>%
  select(antibiotic, pct_change)

adjusted_changes %>%
  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
  geom_bar(stat='identity', fill=NA, color='black') +
  geom_errorbar(width=0.5) +
  geom_point(dat=unadjusted_changes, aes(x=antibiotic, y=pct_change, ymin=NULL, ymax=NULL), color='red') +
  coord_flip() +
  xlab('') +
  ylab('adjusted yearly change in consumption (%)') +
  scale_y_continuous(limits=c(-10, 10), breaks=c(-10, -5, 0, 5, 10)) +
  theme_minimal() +
  theme(axis.title.x = element_text(size=16),
        axis.text = element_text(size=16))
```

@@

adjusted_az_decrease = adjusted_changes %>%
  filter(antibiotic=='azithromycin') %$%
  { (1 + pct_change/100)^3 } %>%
  { (1 - .) * 100 } %>%
  round(2)

adjusted_levo_increase = adjusted_changes %>%
  filter(antibiotic=='levofloxacin') %$%
  { (1 + pct_change/100)^3 } %>%
  { (. - 1) * 100 } %>%
  round(2)
```

@@

read_tsv('tables/model_fq.tsv') %>%
  model_ci() %>%
  select(term, pct_change, ci_low, ci_high, p.value) %>%
  kable
```

Overall, fluoroquinolones changed by about 0% per year during this period. This included `r fq_names`.

@@

I examined encounters recorded in the Carrier Line File whose HCPCS code indicated an office visit. I grouped diagnoses into categories using the Fleming-Dutra eTable 2.

To link diagnoses and PDEs, I looked for all diagnosis-PDE pairs where the PDE came 0-3 days after the diagnosis.

### Delays between diagnosis and prescription

The table shows the number of diagnoses that were followed by that number of days' delay before the following antibiotic prescription. Most prescriptions were filled on the same day as the diagnosis.

```{r dx_delay}
dx_delay = read_tsv('tables/dx_delay.tsv')

dx_delay %>%
  filter(!is.na(delay)) %>%
  mutate(f=n/sum(n)) %>%
  kable
```

@@

### Unadjusted trends in diagnoses

@@

```{r dx_trends}
dx_unadj = read_tsv('tables/dx_counts.tsv') %>%
  select(diagnosis_category, year, dxpkp) %>%
  filter(year %in% c(2011, 2014)) %>%
  spread(year, dxpkp) %>%
  mutate(unadj_pct_change=(`2014`/`2011`-1)*100) %>%
  arrange(desc(`2011`)) %>%
  mutate_if(is.numeric, function(x) round(x, 2))

dx_adj = read_tsv('tables/dx_models.tsv') %>%
  filter(term=='year0') %>%
  mutate(adj_pct_change=100*(exp(estimate*3)-1),
         p.value=p.adjust(p.value, method='BH')) %>%
  mutate_if(is.numeric, function(x) round(x, 2)) %>%
  select(diagnosis_category, adj_pct_change, p.value)

full_join(dx_unadj, dx_adj, by='diagnosis_category') %>%
  kable
```

@@

The first table includes a "no diagnosis" category. This means that the majority of PDEs were not downstream of a diagnosis. In the second table, I remove the "none" diagnosis and rescale the values, which helps make it easier to read.

```{r pde_dx}
read_tsv('tables/pde_dx_counts.tsv') %>%
  filter(antibiotic %in% top_abx, year==2011) %>%
  mutate(pct=round(f_pde_with_dx*100, 1)) %>%
  select(antibiotic, diagnosis_category, pct) %>%
  spread(diagnosis_category, pct) %>%
  kable

read_tsv('tables/pde_dx_counts.tsv') %>%
  filter(antibiotic %in% top_abx, year==2011, diagnosis_category != 'none') %>%
  group_by(antibiotic) %>%
  mutate(f=f_pde_with_dx/sum(f_pde_with_dx),
         pct=round(f*100, 1)) %>%
  ungroup() %>%
  select(antibiotic, diagnosis_category, pct) %>%
  spread(diagnosis_category, pct) %>%
  kable
```

I also did logistic regressions for these trends, but I don't think it's very important. We're not particulary interested in whether the diagnoses that contribute to a drug are changing. We're more interested in whether the drugs that *follow* diagnoses are changing, which is next.

@@

@@

```{r dx_pde_trends}
dx_pde_unadj = read_tsv('tables/dx_pde_counts.tsv') %>%
  filter(antibiotic %in% top_abx, year %in% c(2011, 2014)) %>%
  mutate(key=str_c('y', year)) %>%
  select(diagnosis_category, antibiotic, key, value=f_dx_with_pde)

dx_pde_adj = read_tsv('tables/dx_pde_model.tsv') %>%
  filter(term=='year0') %>%
  mutate(key='OR',
         p.value=p.adjust(p.value, method='BH'),
         OR=round(exp(3*estimate), 2)) %>%
  filter(p.value < 0.05) %>%
  select(diagnosis_category, antibiotic, key, value=OR)

# risk ratio
rr = function(odds_ratio, p0) odds_ratio / (1 - p0 + (p0 * odds_ratio))

dx_pde_both= bind_rows(dx_pde_unadj, dx_pde_adj) %>%
   spread(key, value) %>%
   mutate(pct_change=(y2014/y2011-1)*100,
          adj_pct_change=(rr(OR, y2011)-1)*100) %>%
   select(-OR) %>%
   #mutate_if(is.numeric, function(x) round(x, 2)) %>%
   gather('key', 'value', y2011:adj_pct_change) %>%
   mutate(value=round(value, 2)) %>%
   spread(diagnosis_category, value) %>%
   mutate(key=factor(key, levels=c('y2011', 'y2014', 'pct_change', 'adj_pct_change'))) %>%
   arrange(antibiotic, key)

dx_pde_both %>%
  filter(str_detect(antibiotic, '^a')) %>%
   kable

dx_pde_both %>%
  filter(str_detect(antibiotic, '^c')) %>%
   kable

dx_pde_both %>%
  filter(str_detect(antibiotic, '^[^ac]')) %>%
   kable
```

>>>>

est2pct = function(x) (exp(x) - 1)*100
mult_est = function(df, m) {
  mutate(df, estimate=m*estimate, std.error=sqrt(m)*estimate)
}
model_ci = function(df) {
  df %>%
    mutate(pct_change=est2pct(estimate),
           ci=1.96*std.error,
           ci_low=est2pct(estimate - ci),
           ci_high=est2pct(estimate + ci),
           change_str=sprintf('%.2f%% (%.2f -- %.2f%%)', pct_change, ci_low, ci_high))
}

@@

@@

```{r claims_by_age}
change_demog = read_tsv('tables/model_overall_demography.tsv') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  model_ci() %>%
  filter(term=='year') %>%
  select(name, adj_change=change_str)

@@

```{r adjusted_changes}
adjusted_changes = read_tsv('tables/model_abx.tsv') %>%
  rename(antibiotic=name) %>%
  filter(term=='year') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  model_ci() %>%
  select(antibiotic, pct_change, ci_low, ci_high) %>%
  arrange(pct_change) %>%
  mutate(antibiotic=fct_inorder(antibiotic),
         antibiotic=fct_relevel(antibiotic, 'overall', after=Inf))

unadjusted_changes = bind_rows(claims_by_abx, claims_overall) %>%
  filter(antibiotic %in% c('overall', top_abx), year %in% c(2011, 2014)) %>%
  select(antibiotic, year, cpkp) %>%
  spread(year, cpkp) %>%
  mutate(change=`2014`-`2011`,
         fold_change=change/`2011`,
         pct_change=100*fold_change) %>%
  select(antibiotic, pct_change)

adjusted_changes %>%
  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
  geom_bar(stat='identity', fill=NA, color='black') +
  geom_errorbar(width=0.5) +
  geom_point(dat=unadjusted_changes, aes(x=antibiotic, y=pct_change, ymin=NULL, ymax=NULL), color='red') +
  coord_flip() +
  xlab('') +
  ylab('adjusted change in consumption 2011 to 2014 (%)') +
  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
  theme_minimal() +
  theme(axis.title.x = element_text(size=16),
        axis.text = element_text(size=16))
```

@@

adjusted_az_decrease = adjusted_changes %>%
  filter(antibiotic=='azithromycin') %>%
  pull(pct_change) %>%
  round(2)

adjusted_levo_increase = adjusted_changes %>%
  filter(antibiotic=='levofloxacin') %>%
  pull(pct_change) %>%
  round(2)
```

@@

fq_trends = read_tsv('tables/model_fq.tsv') %>%
  mult_est(3) %>%
  model_ci() %>%
  select(term, pct_change, ci_low, ci_high, p.value) %T>%
  kable

fq_change_pct = fq_trends %>%
  filter(term=='year') %>%
  pull(pct_change) %>%
  round(2)
```

Overall, fluoroquinolones changed by a tiny amount, `r fq_change_pct`%, over this period. This included `r fq_names`.

@@

First, I looked at all encounters recorded in the Carrier Line File and the Outpatient file. A large fraction, I think 78%, of PDEs had an encounter 0-7 days before.

Next, I limited by analysis to encounters who HCPCS code indicated an E&M visit, and I grouped diagnostic codes into categories using Fleming-Dutra eTable 2. I determined the trends in the prevalence of these diagnosis categories by counting the diagnoses per beneficiary per year and by using Poisson regressions that account for the unusual covariates.

I analyzed trends in prescribing practice by computing the fraction of encounters in each diagnosis category that were followed by PDEs for each drug. I computed adjusted trends using logistic regression.

Finally, I considered only those PDEs that are downstream of those E&M diagnoses. For each antibiotic, I computed the fraction of PDEs that have each diagnosis category upstream.

@@

@@

```{r dx_trends}
dx_unadj = read_tsv('tables/dx_trends_table.tsv') %>%
  filter(year %in% c(2011, 2014)) %>%
  left_join(select(totals, year, n_bene), by='year') %>%
  mutate(dxpkp=1000*n_dx/n_bene) %>%
  select(dx_cat, year, dxpkp) %>%
  spread(year, dxpkp) %>%
  mutate(unadj_pct_change=(`2014`/`2011`-1)*100) %>%
  arrange(desc(`2011`)) %>%
  mutate_if(is.numeric, function(x) round(x, 2))

dx_adj = read_tsv('tables/dx_trends.tsv') %>%
  rename(dx_cat=model) %>%
  filter(term=='year') %>%
  mult_est(3) %>%
  model_ci() %>%
  select(dx_cat, change_str, p.value)

full_join(dx_unadj, dx_adj, by='dx_cat') %>%
  kable
```

@@

```{r pde_dx}
read_tsv('tables/dx_from_pde.tsv') %>% 
  spread(present, n) %>%
  mutate(pct=`TRUE`/(`TRUE` + `FALSE`)*100 %>% round(2)) %>%
  select(antibiotic, dx_cat, pct) %>%
  spread(dx_cat, pct) %>%
  head %>% kable
```

@@

```{r dx_to_pde_table}
dx_to_pde_unadj = read_tsv('tables/dx_rx_table.tsv') %>%
  filter(year %in% c(2011, 2014)) %>%
  select(dx_cat, antibiotic, year, f) %>%
  spread(year, f)

dx_to_pde_adj = read_tsv('tables/dx_rx_trends.tsv') %>%
  filter(term=='year') %>%
  mult_est(3) %>%
  mutate(ci=1.96*std.error,
         ci_low=estimate-ci,
         ci_high=estimate+ci) %>%
  select(dx_cat, antibiotic, estimate, ci_low, ci_high)

risk_ratio_f = function(or, p) or / (1 + p * (or - 1))

inner_join(dx_to_pde_unadj, dx_to_pde_adj, by=c('dx_cat', 'antibiotic')) %>%
  mutate(estimate=risk_ratio_f(estimate, `2011`),
         ci_low=risk_ratio_f(ci_low, `2011`),
         ci_high=risk_ratio_f(ci_high, `2011`),
         pct_change=(`2014`-`2011`)/`2011`) %>%
  mutate_if(is.numeric, function(x) round(x*100, 2)) %>%
  mutate(adj_change=sprintf('%.2f%% (%.2f -- %.2f%%)', estimate, ci_low, ci_high)) %>%
  select(dx_cat, antibiotic, `2011`, `2014`, pct_change, adj_change) %>%
  kable
```

@@

```{r dx_to_pde_trends}

```
